---
title: "Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(p8105.datasets)
library(dplyr)
library(plotly)
library(readxl)
```
Import New York state weather stations data from January 1, 1981 through December 31, 2010.
```{r include=FALSE}
nynoaadat <- read_csv("nynoaadat.csv")
nynoaadat_cleaned <- nynoaadat %>%
  filter(!(is.na(prcp) & is.na(snow) & is.na(snwd) & is.na(tmax) & is.na(tmin)))
view(nynoaadat_cleaned)
```
The data of NY  state weather stations data shows the 
id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C) from January 1, 1981 through December 31, 2010. It contains `r nrow(nynoaadat_cleaned)` observation and `r ncol(nynoaadat_cleaned)` columns.

Mutate the cleaned data to prepare for making plots
```{r include=FALSE}
nynoaadat_cleaned <- nynoaadat %>%
  mutate(
    date = as.Date(date),
    month = month(date, label = TRUE),
    year = year(date),
    tmax = tmax / 10,  # Convert to degrees C
    tmin = tmin / 10,  # Convert to degrees C
    prcp = prcp / 10   # Convert to mm
  )
```

```{r include=FALSE}
# Create station summary for scatter plot
station_summary <- nynoaadat_cleaned %>%
  group_by(id) %>%
  summarize(
    avg_tmax = mean(tmax, na.rm = TRUE),
    avg_prcp = mean(prcp, na.rm = TRUE),
    n_obs = n()
  ) %>%
  filter(!is.na(avg_tmax), !is.na(avg_prcp))
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}
nynoaadat_cleaned %>%
  plot_ly(
    x = ~month,
    y = ~tmax,
    type = "box",
    name = "Max Temp",
    color = I("#1f77b4")
  ) %>%
  add_trace(
    y = ~tmin,
    name = "Min Temp",
    color = I("#ff7f0e")
  ) %>%
  layout(
    title = "Temperature Distribution by Month",
    xaxis = list(title = "Month"),
    yaxis = list(title = "Temperature (°C)"),
    showlegend = TRUE
  )
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}
plot_ly(
  data = station_summary,
  x = ~avg_tmax,
  y = ~avg_prcp,
  type = "scatter",
  mode = "markers",
  marker = list(
    size = 8,
    opacity = 0.6
  ),
  color = ~avg_tmax,
  text = ~paste(
    "Station ID:", id,
    "\nAvg Max Temp:", round(avg_tmax, 1), "°C",
    "\nAvg Precipitation:", round(avg_prcp, 1), "mm",
    "\nObservations:", n_obs
  )
) %>%
  layout(
    title = "Average Maximum Temperature vs Precipitation by Station",
    xaxis = list(title = "Average Maximum Temperature (°C)"),
    yaxis = list(title = "Average Precipitation (mm)"),
    hovermode = "closest"
  )
```

### Chart C

```{r}
nynoaadat_cleaned %>%
  group_by(year) %>%
  summarize(
    avg_snow = mean(snow, na.rm = TRUE)
  ) %>%
  filter(!is.na(avg_snow)) %>%
  plot_ly(
    x = ~year,
    y = ~avg_snow,
    type = "bar",
    marker = list(
      color = ~avg_snow,
      colorscale = "Viridis"
    ),
    text = ~paste(
      "Year:", year,
      "\nAvg Snowfall:", round(avg_snow, 1), "mm"
    )
  ) %>%
  layout(
    title = "Average Annual Snowfall",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Average Snowfall (mm)"),
    showlegend = FALSE
  )
```

